---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data

1. Load the data (i.e. read.csv())

```{r}
unzip("activity.zip")
raw <- read.table("activity.csv", sep = ",", header = TRUE, na.strings = "NA")
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
library(data.table)
data <- raw
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data <- data[complete.cases(data),]
data <- data.table(data)
summary(data)
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
totalStepsPerDay <- data[,sum(steps),by=date]
```

2. Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
ggplot(totalStepsPerDay, aes(x = V1)) +
  geom_histogram(stat="bin", binwidth=4000) +
  xlab("Total steps for each day")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(totalStepsPerDay$V1)
median(totalStepsPerDay$V1)
```


## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
meanStepsPerInterval <- data[,mean(steps),by=interval]

ggplot(meanStepsPerInterval, aes(x = interval, y = V1)) +
  geom_line() +
  xlab("Interval") +
  ylab("Mean number of steps for each day")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
meanStepsPerInterval[which.max(meanStepsPerInterval$V1)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(raw))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
dataWithNAs <- raw
dataWithNAs$date <- as.Date(dataWithNAs$date, format = "%Y-%m-%d")
dataWithNAs <- data.table(dataWithNAs)

meanStepsPerInterval <- data.table(meanStepsPerInterval)

# For every row with "NA", apply the Avg Steps for the "interval" in that row.
# The Avg Steps is a double number, so its converted to integer.
naIndex = (1:nrow(raw))[is.na(raw)]
dataWithNAs[naIndex,1] <- as.integer(
  dataWithNAs[naIndex,
              meanStepsPerInterval[interval == interval]]$V1
  )
```


3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
summary(dataWithNAs)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
totalStepsPerDay <- dataWithNAs[,sum(steps),by=date]

ggplot(totalStepsPerDay, aes(x = V1)) +
  geom_histogram(stat="bin", binwidth=4000) +
  xlab("Total steps for each day")

mean(totalStepsPerDay$V1)
median(totalStepsPerDay$V1)
```

* Q: Do these values differ from the estimates from the first part of the assignment?
     * A: They do differ, but not by much.
* Q: What is the impact of imputing missing data on the estimates of the total daily number of steps?
     * A: The data seems to be slightly more concentraded near the mean.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
data$weekend <- lapply(weekdays(data$date), function(x) x == "Sunday" || x == "Saturday")
data$weekend <- factor(data$weekend == TRUE)
```

   1. Number os weekends and weekdays

```{r}
summary(data$weekend)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
weekendIndex <- (1:nrow(data))[data$weekend == TRUE]
weekdayIndex <- (1:nrow(data))[data$weekend == FALSE]
meanStepsWeekend <- data[weekendIndex,mean(steps),by=interval]
meanStepsWeekday <- data[weekdayIndex,mean(steps),by=interval]

meanStepsWeekend$weekend <- TRUE
meanStepsWeekday$weekend <- FALSE

meanSteps <- rbind(meanStepsWeekday, meanStepsWeekend)

ggplot(meanSteps, aes(x = interval, y = V1, col = weekend)) +
  geom_line() +
  xlab("interval") +
  ylab("Mean steps")
```